\papertitle

### 2023-06-09 src/tex/ms.tex
Matchign v3 on arXiv

% Import relevant packages
\usepackage{showyourwork}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{multirow}
\usepackage[separate-uncertainty=true]{siunitx}
\usepackage{amsmath}
\usepackage[shortlabels]{enumitem}
\usepackage[page]{appendix}
\usepackage{xfrac}
\usepackage{printlen}
\usepackage[version=4]{mhchem}
\usepackage{blindtext}
% \usepackage[hidelinks]{hyperref}
\sisetup{
    list-units      = brackets,
    range-units     = brackets,
    range-phrase    = -,
    list-pair-separator= {, },
    list-separator  = {, },
    list-final-separator = {, }
    }
\renewcommand\appendixname{}
\newcommand{\papertitle}{Accurate Estimation of Diffusion Coefficients and their Uncertainties from Computer Simulation}
% \newcommand{\oMSD}{\ensuremath{\overline{\bm{s}^2}}}
% \newcommand{\oMSDs}[1]{\ensuremath{\overline{s^2}}(#1)}
% \newcommand{\oMSDi}{\ensuremath{\overline{s^2}_i}}
% \newcommand{\oMSDj}{\ensuremath{\overline{s^2}_j}}

@@

\let\oldaddcontentsline\addcontentsline% Store \addcontentsline
\renewcommand{\addcontentsline}[3]{}

@@

For data that are both statistically independent and identically normally distributed, ordinary least squares regression (OLS) is unbiased and statistically efficient, and gives accurate estimates of the uncertainties in the resulting regression coefficients.
MSD data obtained from simulation, however, are neither statistically independent nor identically distributed.
The variances, $\var{\oMSDi}$, are correlated, since the displacement of each particle at time $t+\Delta t$ is necessarily similar to its displacement at time $t$, and hence, $\oMSDs{t}$ is similar to $\oMSDs{t+\Delta t}$.
These variances are also typically unequal---the data are heteroscedastic;
for example, for a random walk, the MSD variance increases quadratically with time \cite{smith_random_1996}. 
Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (Fig.~\ref{fig:glswlsols}a).

@@

A second method for estimating the ensemble-average MSD, and thus $\Dest$, from simulation data is Bayesian regression.
Like GLS, Bayesian regression can take into account both the changing variance and the correlation structure inherent in the data.
Rather than providing a singular ``best-fit'' estimate like GLS, Bayesian regression produces a posterior probability distribution for the regression coefficients.
The mean of this distribution serves as the point estimate of the coefficients and, in the absence of additional prior information, is equivalent to the GLS estimate, while the spread of the distribution quantifies the uncertainty in these estimates.
For data that is both heteroscedastic and correlated, such as MSD data from simulations, Bayesian regression, like GLS, is formally fully statistically efficient.

@@

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (Fig.~\ref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (Fig.~\ref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via Eqn.~\ref{equ:cvv}.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/msd.pdf}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}
%

@@

While the analytical and average estimated covariance matrices show some systematic deviation from the numerically converged covariance matrix, the general correlation structure is preserved.
The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the component covariance matrices estimated from individual simulation trajectories may be used within approximte GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.

@@

To demonstrate the complete approximate Bayesian regression scheme, as described above, we present two distinct examples.
First, we consider a simple 3D-lattice random walk, where the true self-diffusion coefficient $\D$ is specified by the simulation parameters, and a well-converged numerical covariance matrix can be obtained with relatively low computational cost, which allows us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
Second, we consider an example real-world system---the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO)---which represents a practical application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.

@@

@@

We next consider the degree to which our method can quantify the uncertainty in $\Dest$ when using input data from a single simulation.
Fig.~\ref{fig:random_walk}d shows the distribution of estimated variances $\varest{\Dest}$, with each sample calculated from an individual simulation trajectory.
We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$.
In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
Notably, the precision of this estimate is significantly greater than obtained using OLS or WLS and their corresponding textbook statistical formulae.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/diffusion.pdf}
    \caption{
      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    }
    \label{fig:diffusion}
    \script{diffusion.py}
\end{figure}

@@

@@

\title{Supplemental Material for ``\papertitle''}
\maketitle
% \onecolumngrid

This document presents supplementary material for the manuscript ``\papertitle''.
It contains the following sections:
\begin{enumerate}
    % \item Evidence for the effect of the number of independent observations on the distribution of mean-squared displacements.
    \item The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
    \item Analysis of the relative efficiency of the approximate Bayesian regression method, presented here, compared to OLS, WLS, and GLS, and the efficiency scaling of each method with number of mobile particles or total simulation time..
    % \item Discussion of the skewed distribution of the estimated variance in the diffusion coefficient.
    \item Details of the LLZO molecular dynamics simulations.
    % \item Discussion of the importance of only using non-diffusive simulation in the analysis.
    \item Details of the implementation of the model covariance method in the \textsc{kinisi} package.
\end{enumerate}
A repository containing the analysis and plotting code used to generate all results and figures in the main manuscript and this supplemental material document is available at \url{www.github.com/arm61/msd-errors}~\cite{mccluskey_github_2022}, under MIT (code) and CC BY-SA 4.0 (figures and text) licenses.
This repository includes a fully reproducible \code{showyourwork} workflow, which allows complete reproduction of the analysis, plotting of figures and compilation of the manuscripts.
The corresponding input datasets are openly available under the CC BY-SA 4.0 licence \cite{mccluskey_zenodo_2022,coles_llzo_zenodo_2022}.

% \vspace{\columnsep}
\twocolumngrid

@@

Eqn.~\ref{equ:varmsd} gives the variance of the mean squared displacement for a single particle considering a single time-origin.
We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be acheived by averaging over mobile particles or by averaging over time origins.
This averaging over equivalent observations reduces the variance in the observed MSD to
%
\begin{equation}
    \var{\oMSDn} = \frac{2n^2\kappa^4}{\nind{n}},
\label{equ:der_var}
\end{equation}
%
where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of non-overlapping time-windows of length $i$ in our simulation trajectory.
Note that $\nind{n}$ considers non-overlapping time windows, since mutually overlapping time-windows give correlated squared displacements.

The results for a one-dimensional lattice above (Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var}) can be extended to a $d$-dimensional lattice, to give
%
\begin{equation}
    \moMSDn_{d} = \sum^d{\frac{n\kappa^2}{d}} = n\kappa^2,
\end{equation}
%
with variance
%
\begin{equation}
    \var{\oMSDn}_d = \sum^d{\frac{2n^2\kappa^4}{d^2\nind{n}}} = \frac{2n^2\kappa^4}{d\nind{n}},
\end{equation}
%.
Because each step is equally likely to move a partice along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.

The analysis above can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
%
\begin{equation}
    \begin{aligned}
        \moMSDn &= n\kappa^2; \\
        \moMSDm &= (n+m)\kappa^2.
    \end{aligned}
\end{equation}
%
The covariance between these is defined as
%
\begin{equation}
    \Sigma \left[\oMSDn, \oMSDm \right] = \left<{\left[{\oMSDn - \moMSDn}\right] \left[{\oMSDm - \moMSDm}\right]}\right>, 
\end{equation}
%
which can be expanded as
%
\begin{equation}
    \begin{aligned}
        \Sigma \left[\oMSDn, \oMSDm \right] = \big\langle & \oMSDn\oMSDm - \oMSDn\moMSDm \\
        & - \moMSDn\oMSDm + \moMSDn\moMSDm\big\rangle,
    \end{aligned}
\end{equation}
%
and then reformulated to give
%
\begin{equation}
    \Sigma \left[\oMSDn, \oMSDm \right] = \left<\oMSDn\oMSDm\right> - \moMSDn\moMSDm,
    \label{equ:pair}
\end{equation}
%
where 
%
\begin{equation}
    \begin{aligned}
        \moMSDn\moMSDm &= \oMSDn\oMSDm \\
                       &= nd^2(n+m)\kappa^2 \\
                       &= n(n+m)\kappa^4
    \end{aligned}
\end{equation}
%
and, by analogy to Eqn.~\ref{equ:big_av},
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = \left<\sum_i^n\sum_j^n\sum_k^{n+m}\sum_l^{n+m} h_i h_j h_k h_l\right>,
\end{equation}
%
which we can rewrite as
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> = \Bigg\langle & \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=1}^{n} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
    \end{aligned}
    \label{equ:vbig}
\end{equation}
%
The second and third terms in Eqn.~\ref{equ:vbig} tend to \num{0} as there is an equal probability of positive and negative displacements. 
This reduces Eqn.~\ref{equ:vbig} to 
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> = & \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \Bigg\rangle \\
        & + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle, 
    \end{aligned}
\end{equation}
%
and using Eqn.~\ref{equ:infty} gives
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
\end{equation}
%
We can rewrite this as
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n h_i h_j \Bigg\rangle \Bigg\langle \sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_k h_l \Bigg\rangle,
\end{equation}
%
where the following holds,
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> &= 3n^2\kappa^4 + n\kappa^2 m\kappa^2 \\
                                  &= 3n\kappa^4 + nm\kappa^4.
    \end{aligned}
\end{equation}
%
Putting this result into Eqn.~\ref{equ:pair} allows the covariance to be written as
%
\begin{equation}
    \begin{aligned}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] &= 3n^2\kappa^4 + nm\kappa^4 - n(n+m)\kappa^4 \\
                                        &= 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
    \end{aligned}
    \label{equ:cov_der}
\end{equation}
where we use the $\Sigma^\prime$ notation to identify that this is in the long-time limit.

In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
We can easily rationalise this by accepting that for a random walk any non-overlapping point will be completely uncorrelated and therefore have a covariance of \num{0}. 
Similar to the case for the variance, the covariance derived in Eqn.~\ref{equ:cov_der} is that for a single particle at a single time origin. 
The number of independent observed squared displacements for a given covariance should be the minimum number of shared independent observed squared displacements between the two time intervals, which is $\nind{n+m}$. 
Therefore, scaled by the number of contributing independent observations, the covariance, in the long-time limit, is
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{\nind{n+m}}.
\end{equation}
%
Similar to the MSD and the variance, the covariance can be written for $d$-dimensions as 
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{d\nind{n+m}}.
\end{equation}
%
The covariance can be calculated directly from the variance by recognising that both depend on the number of overlapping points, $n$, as follows
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \var{\oMSDn}\frac{\nind{n}}{\nind{n+m}}.
\end{equation}
%
This is then rewritten in terms of $i$ and $j$ to give, Eqn.~\ref{equ:cvv_SI}.

>>>>

% Import relevant packages
\usepackage{showyourwork}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{multirow}
\usepackage[separate-uncertainty=true]{siunitx}
\usepackage{amsmath}
\usepackage[shortlabels]{enumitem}
\usepackage[page]{appendix}
\usepackage{xfrac}
\usepackage{printlen}
\usepackage[version=4]{mhchem}
\usepackage{blindtext}
\sisetup{
    list-units      = brackets,
    range-units     = brackets,
    range-phrase    = -,
    list-pair-separator= {, },
    list-separator  = {, },
    list-final-separator = {, }
    }
\renewcommand\appendixname{}
\newcommand{\papertitle}{Accurate Estimation of Diffusion Coefficients and their Uncertainties from Computer Simulation}

@@

\let\oldaddcontentsline\addcontentsline
\renewcommand{\addcontentsline}[3]{}

@@

For data that are both statistically independent and identically normally distributed, ordinary least squares regression (OLS) is unbiased and statistically efficient, and gives accurate estimates of the uncertainties in the resulting regression coefficients.
MSD data obtained from simulation, however, are neither statistically independent nor identically distributed.
The variances, $\var{\oMSDi}$, are correlated, since the displacement of each particle at time $t+\Delta t$ is necessarily similar to its displacement at time $t$, and hence, $\oMSDs{t}$ is similar to $\oMSDs{t+\Delta t}$.
These variances are also typically unequal---the data are heteroscedastic \cite{smith_random_1996,he_statistical_2018}. 
Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (Fig.~\ref{fig:glswlsols}a).

@@

An alternative method for estimating the ensemble-average MSD, and thus $\Dest$, from simulation data is Bayesian regression.
Like GLS, Bayesian regression can take into account both the changing variance and the correlation structure inherent in the data.
Rather than providing a singular ``best-fit'' estimate like GLS, Bayesian regression produces a posterior probability distribution for the regression coefficients.
The mean of this distribution serves as the point estimate of the coefficients and, in the absence of additional prior information, is equivalent to the GLS estimate, while the spread of the distribution quantifies the uncertainty in these estimates.
For data that is both heteroscedastic and correlated, such as MSD data from simulations, Bayesian regression, like GLS, is formally fully statistically efficient.

@@

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (Fig.~\ref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (Fig.~\ref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via Eqn.~\ref{equ:cvv}.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/msd.pdf}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}
%

@@

While the analytical and average estimated covariance matrices show some systematic deviation from the numerically converged covariance matrix, the general correlation structure is preserved.
The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the covariance matrices estimated from individual simulation trajectories may be used within approximate GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.

@@

@@

To demonstrate the complete approximate Bayesian regression scheme, as described above, we present two distinct examples.
First, we consider a simple 3D-lattice random walk, where the true self-diffusion coefficient $\D$ is specified by the simulation parameters, and a well-converged numerical covariance matrix can be obtained with relatively low computational cost, which allows us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
Second, we consider an example real-world system---the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO)---which represents an application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.

@@

We next consider the degree to which our method can quantify the uncertainty in $\Dest$ when using input data from a single simulation.
Fig.~\ref{fig:random_walk}d shows the distribution of estimated variances $\varest{\Dest}$, with each sample calculated from an individual simulation trajectory.
We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$ (further details are provided in the SI).
In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
Notably, the precision of this estimate is significantly greater than obtained using OLS or WLS and their corresponding textbook statistical formulae.

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{figures/diffusion.pdf}
    \caption{
      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    }
    \label{fig:diffusion}
    \script{diffusion.py}
\end{figure}

@@

\section{$\var{\Dest}$ scaling and comparison to OLS, WLS, and GLS}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{figures/stateff.pdf}
    \caption{
      Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
      (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
      (b) Scaling versus total simulation time, $t_\mathrm{max}$.
      Solid lines show fitted power law relationships for each dataset.
      The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
    }
    \label{fig:stat_eff}
    \script{stat_eff.py}
\end{figure}

Fig.~\ref{fig:stat_eff} presents an analysis of the variation in $\var{\Dest}$ as the number of mobile particles (Fig.~\ref{fig:stat_eff}a) and the total simulation time (number of steps) (Fig.~\ref{fig:stat_eff}b) are changed.
We compare four methods for estimating $\D$ from the observed MSD data: OLS, WLS, the approximate Bayesian regression method described here, and GLS.
When estimating $\D$ using WLS and GLS, we calculate the variances and the covariance matrix, respectively, numerically, using the complete set of \num{512} simulations.
Each data point in Fig.~\ref{fig:stat_eff} represents the variance across point-estimates of $\D$ derived from \num{512} individual 3D-lattice random walk simulations, for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
The GLS dataset corresponds to an optimally efficient estimator for linear regression of observed MSD data, equivalent to performing Bayesian regression with the numerical covariance matrix and an uninformative prior.

Our approximate Bayesian regression method performs similarly to GLS, with a numerically converged covariance matrix, and gives significantly reduced uncertainty in $\Dest$ compared to OLS or WLS, for all simulation sizes and lengths considered.
Moreover, our method scales better than OLS or WLS as the total simulation time is increased.
The approximate Bayesian regression method in this manuscript therefore presents a significant improvement over more conventional methods such as OLS and WLS, by enabling more precise estimates of $\D$ across varied simulation sizes at equivalent computational cost.

@@

\title{Supplemental Material for ``\papertitle''}
\maketitle

This document presents supplementary material for the manuscript ``\papertitle''.
It contains the following sections:
\begin{enumerate}
    \item The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
    \item Discussion of the origin of bias and skew in the distribution of the estimated variance of the estimated diffusion coefficient, $\prob{\varest{\Dest}}$.
    \item Details of the LLZO molecular dynamics simulations.
    \item Details of the implementation of the model covariance method in the \textsc{kinisi} package.
\end{enumerate}
A repository containing the analysis and plotting code used to generate all results and figures in the main manuscript and this supplemental material document is available at \url{www.github.com/arm61/msd-errors}~\cite{mccluskey_github_2022}, under MIT (code) and CC BY-SA 4.0 (figures and text) licenses.
This repository includes a fully reproducible \code{showyourwork} workflow, which allows complete reproduction of the analysis, plotting of figures and compilation of the manuscripts.
The corresponding input datasets are openly available under the CC BY-SA 4.0 licence \cite{mccluskey_zenodo_2022,coles_llzo_zenodo_2022}.

\twocolumngrid

@@

Eqn.~\ref{equ:varmsd} gives the variance of the mean squared displacement for a single particle considering a single time-origin.
We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be achieved by averaging over mobile particles or by averaging over time origins.
This averaging over equivalent observations reduces the variance in the observed MSD to
%
\begin{equation}
    \var{\oMSDn} = \frac{2n^2\kappa^4}{\nind{n}},
\label{equ:der_var}
\end{equation}
%
where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of non-overlapping time-windows of length $i$ in our simulation trajectory.
Note that $\nind{n}$ considers non-overlapping time windows, since mutually overlapping time-windows give correlated squared displacements.

The results for a one-dimensional lattice above (Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var}) can be extended to a $d$-dimensional lattice, to give
%
\begin{equation}
    \moMSDn_{d} = \sum^d{\frac{n\kappa^2}{d}} = n\kappa^2,
\end{equation}
%
with variance
%
\begin{equation}
    \var{\oMSDn}_d = \sum^d{\frac{2n^2\kappa^4}{d^2\nind{n}}} = \frac{2n^2\kappa^4}{d\nind{n}},
\end{equation}
%.
Because each step is equally likely to move a particle along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.

The analysis above can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
%
\begin{equation}
    \begin{aligned}
        \moMSDn &= n\kappa^2; \\
        \moMSDm &= (n+m)\kappa^2.
    \end{aligned}
\end{equation}
%
The covariance between these is defined as
%
\begin{equation}
    \Sigma \left[\oMSDn, \oMSDm \right] = \left<{\left[{\oMSDn - \moMSDn}\right] \left[{\oMSDm - \moMSDm}\right]}\right>, 
\end{equation}
%
which can be expanded as
%
\begin{equation}
    \begin{aligned}
        \Sigma \left[\oMSDn, \oMSDm \right] = \big\langle & \oMSDn\oMSDm - \oMSDn\moMSDm \\
        & - \moMSDn\oMSDm + \moMSDn\moMSDm\big\rangle,
    \end{aligned}
\end{equation}
%
and then reformulated to give
%
\begin{equation}
    \Sigma \left[\oMSDn, \oMSDm \right] = \left<\oMSDn\oMSDm\right> - \moMSDn\moMSDm,
    \label{equ:pair}
\end{equation}
%
where 
%
\begin{equation}
    \begin{aligned}
        \moMSDn\moMSDm &= \oMSDn\oMSDm \\
                       &= n\kappa^2(n+m)\kappa^2 \\
                       &= n(n+m)\kappa^4
    \end{aligned}
\end{equation}
%
and, by analogy to Eqn.~\ref{equ:big_av},
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = \left<\sum_i^n\sum_j^n\sum_k^{n+m}\sum_l^{n+m} h_i h_j h_k h_l\right>,
\end{equation}
%
which we can rewrite as
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> = \Bigg\langle & \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=1}^{n} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
    \end{aligned}
    \label{equ:vbig}
\end{equation}
%
The second and third terms in Eqn.~\ref{equ:vbig} tend to zero as there is an equal probability of positive and negative displacements. 
This reduces Eqn.~\ref{equ:vbig} to 
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> = & \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \Bigg\rangle \\
        & + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle, 
    \end{aligned}
\end{equation}
%
and using Eqn.~\ref{equ:infty} gives
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
\end{equation}
%
We can rewrite this as
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n h_i h_j \Bigg\rangle \Bigg\langle \sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_k h_l \Bigg\rangle,
\end{equation}
%
where the following holds,
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> &= 3n^2\kappa^4 + n\kappa^2 m\kappa^2 \\
                                  &= 3n\kappa^4 + nm\kappa^4.
    \end{aligned}
\end{equation}
%
Putting this result into Eqn.~\ref{equ:pair} allows the covariance to be written as
%
\begin{equation}
    \begin{aligned}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] &= 3n^2\kappa^4 + nm\kappa^4 - n(n+m)\kappa^4 \\
                                        &= 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
    \end{aligned}
    \label{equ:cov_der}
\end{equation}
where we use the $\Sigma^\prime$ notation to identify that this is in the long-time limit.

In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
We can rationalise this by noting that for a random walk any non-overlapping points will be completely uncorrelated and therefore have a covariance of \num{0}. 
Similar to the case for the variance, the covariance derived in Eqn.~\ref{equ:cov_der} is that for a single particle at a single time origin. 
The number of independent observed squared displacements for a given covariance should be the minimum number of shared independent observed squared displacements between the two time intervals, which is $\nind{n+m}$. 
Therefore, the covariance, scaled by the number of contributing independent observations, in the long-time limit, is
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{\nind{n+m}}.
\end{equation}
%
Similar to the MSD and the variance, the covariance can be written for $d$-dimensions as 
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{d\nind{n+m}}.
\end{equation}
%
The covariance can be calculated directly from the variance by recognising that both depend on the number of overlapping points, $n$, as follows
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \var{\oMSDn}\frac{\nind{n}}{\nind{n+m}}.
\end{equation}
%
This is then rewritten in terms of $i$ and $j$ to give, Eqn.~\ref{equ:cvv_SI}.
